### 2018-03-16 paper/paper.tex
start paper template

\documentclass[letterpaper]{article}

\input{paper.sty}
% Load nips_2017 last to override any changes I didn't mean to make.
\usepackage[final]{nips_2017}

\title{Generating haikus with an LSTM network}

\author{
    Austin Gill\\
    Department of Mathematics and Computer Science\\
    South Dakota School of Mines and Technology\\
    Rapid City SD, 57701\\
    \texttt{\href{mailto:austin.gill@mines.sdsmt.edu}{austin.gill@mines.sdsmt.edu}} \\
}

\begin{document}
\maketitle

\begin{abstract}
    Here's an abstract.
\end{abstract}

\section{Introduction}
    Here's the introduction.

\section{Stuff}
    Here's some stuff you didn't know!

\section{Conclusion}
    Conclude something\todo{Something useful!}

\end{document}